---
title       : Positivity metric calculator
subtitle    : Course Project for Developing Data Products class
author      : Alexey Ermakov
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## What is it?

* This is attempt to get emotional context for given text
* Is it something positive or negative? 
* The app is trying to estimate this and give numeric result on scale 0 to 10
* 0 means text is totaly negative and you probably shouldn't read it :)
* 10 means the text is awesome

--- .class #id 

## How it works
* The team from University of Vermon made a great research [Human language reveals a universal positivity bias] (http://arxiv.org/pdf/1406.3855v1.pdf). They get around 10000 most frequent english words and ask thousands of native speakers to rate each word on a 9 point scale, with 1 corresponding to most negative or saddest, 5 to neutral, and 9 to most positive or happiest.
* The results were [published] (http://www.uvm.edu/storylab/share/papers/dodds2014a/data.html). For each language they provide 2 files: one with words and one with average rate for each word.
* Shiny application process given text using [Text Mining](http://cran.r-project.org/web/packages/tm/) package removing punctuation, stop words, white spaces and gets matrix of words and their counts.
* Then using files from the research app is calculating average rate of the dictionary words which were in text.
* The result is normalized number from 0 to 10 which represents "positivity" of given text and 2 lists with most positive and most negative words

--- .class #id 

## Example
##### get matrix of words for text from the news
```{r echo=FALSE}
setwd("../")
source("server.R")
```
```{r}
wordMatrix <- getWordMatrix("The leaders of the House and Senate Veterans Affairs committees reached a tentative deal Sunday to improve veterans' health care -- a potential solution to help fix such ongoing problems as delays for benefits and long waits for medical treatment concealed by secret lists.")
wordMatrix[1:5,]
```

```{r}
getPositivity(wordMatrix)
```

--- .class #id 

## Try it by yourself!

* Shiny App [http://alexius.shinyapps.io/text-positivity-metric/](http://alexius.shinyapps.io/text-positivity-metric/)


